Convergence rate of MAP estimates \\ for the exponential family

### 2021-02-23 main.tex
Add table of contents

@@

\subsection{Transferring SMD's covariance assumption to MAP estimates}
In our setting, 
\begin{align}
	\nabla f(\natp_t) - g_t =  T(X_{t+1})	- \E[T(X)]
\end{align}
so that $\sigma$ is really a bound on the covariance of the sufficient statistics with another variable
\begin{align}
	&\Cov( T(X_{t+1}), \natp_{t+1}| X_{1\dots t})
	\leq \lr_t\sigma^2 \\
	=& \Cov(T(X_{t+1}), \nabla\conj \left(\lr_t T(X_{t+1}) + (1-\lr_t ) \mu_t \right) | X_{1\dots t}) \;.
\end{align}
In other words, we need for all $\lr\in(0,\inv{n})$, and $ \mu$ in the interior of the marginal polytope,
\begin{align}
\label{eq:hypvariance}
	\Cov_X \bigg(T(X), \nabla \conj (\lr T(X) + (1-\lr) \mu)\bigg	) 
	\leq \lr \sigma^2 \; .
\end{align}
We study this assumption in Section \ref{sec:gaussian_variance} where we show that it cannot hold in general for the $\mu$ we care about. 
\RLP{should we get a bound that depends on mu instead ? or split between $\mu$ smaller or larger than $\varepsilon$}
When we plug this bound into the expected suboptimality formula~\eqref{eq:bregmanMLE} with the maximum likelihood estimate $\hat \mu =  \inv{n}\sum_i T(X_i)$, assuming all $T(X)$ belong to the marginal polytope, we get
\begin{multline}
	\expect[X_{1\dots n}]{\bregmanconj ( \mu_* ;\hat \mu) }
	=\conj(\mu_*) \overbrace{- \E[\conj(\hat \mu)]}^{\leq - \conj(\E[\hat \mu])} \\
	+ \inv{n}\sum_i \E[\underbrace{\E[\langle  T(X_i) - \mu_*; \nabla\conj (\hat \mu) \rangle | X_j, j\neq i]}_{\leq \nicefrac{\sigma^2}{n} \eqref{eq:hypvariance}
	 	\; \text{with}\; \mu=\inv{n-1}\smallsum_{j\neq i} T(x_j)  } ]  \leq \frac{\sigma^2}{n}
\end{multline}
\textcolor{red}{what about:}
\begin{multline}
	\expect[X_{1\dots n}]{\bregmanconj ( \mu_* ;\hat \mu) }
	=\conj(\mu_*) \overbrace{- \E[\conj(\hat \mu)]}^{\leq - \conj(\E[\hat \mu])} \\
	+ \inv{n}\sum_i \E[\underbrace{\E[\langle  T(X_i) \textcolor{red}{- \hat \mu_{n-1} + \hat \mu_{n-1}} - \mu_*; \nabla\conj (\hat \mu) \rangle | X_j, j\neq i]}_{\leq \nicefrac{\sigma^2}{n} \eqref{eq:hypvariance} } ]  \leq \frac{\sigma^2}{n}
\end{multline}
where we used the decomposition $\hat \mu = \inv{n}T(X_i) + (1-\inv{n}) \inv{n-1} \sum_{j\neq i}T(X_j)$ to apply~\eqref{eq:hypvariance}. 
In words, this variance assumption on $T(X)$ and $\conj$ immediately gives us a bound on the suboptimality. We can also apply this result to the MAP estimate $\hat \mu = \frac{n_0 \mu_0 + \smallsum_i T(X_i)}{n_0+n}$ but I did not manage to reach a satisfying conclusion about the bias. 
Note $\tilde \mu = \E[\hat \mu] = \frac{n_0 \mu_0 + n \mu_*}{n_0+n}$ and $\gamma_0= \frac{n_0}{n_0+n}$ and the step-size $\gamma = \inv{n_0 + n}$ .
\begin{align}
	\expect[X_{1\dots n}]{\bregmanconj ( \mu_* ;\hat \mu) } 
	&\leq \conj(\mu_*) - \conj(\tilde \mu) + \gamma_0 \langle \mu_0 - \mu_* ; \E[\hat \natp] \rangle + (1 - \gamma_0) \gamma \sigma^2  \\
	& = \bregmanconj (\mu_* ; \tilde \mu) - \langle \tilde \mu - \mu_* ;    \tilde \natp - \E[\hat \natp] \rangle + \frac{n}{(n_0+n)^2}\sigma^2
\end{align}
so we recover the $O(n^{-1})$ for the variance and we are still looking for $O(n^{-2})$ rate for the bias. The Bregman term in the bias should be in this spirit, given a quadratic approximation
\begin{align}
	 \bregmanconj (\mu_* ; \tilde \mu)  
	 = \bregmanconj (\mu_* ; \mu_* + \gamma_0 (\mu_0 - \mu_*) ) \approx \half \gamma_0^2 \norm{\mu_0 - \mu_*}^2_{\mSigma_*^{-1}}
\end{align}
where $\mSigma_* = \Cov_{\mu_*}(T(X))$. 
The order of terms $\bregmanconj (\mu_* ; \tilde \mu)  \neq \bregmanconj (\tilde \mu ; \mu_* ) $ only affects third order terms.
That's for the bregman, but the scalar product is harder to bound. Hopefully it could be positive. If not we know that $\tilde \mu - \mu_* = \frac{n_0}{n_0+n} (\mu_0 - \mu_*) = O(n^{-1})$ , and we can hope that the same holds for $\tilde \natp - \E[\hat \natp] = \nabla \conj(\E[\hat \mu]) - \E[\nabla\conj (\hat \mu)]$, which kinda measures the non-linearity, or the curvature of $\conj$. Using a simple cauchy-schwartz, and the hessian of the conjugate, we might be able to get something.

@@

\paragraph{Open questions}
The question remains of : when is such a bound valid?  what does $\sigma^2$ depend on ? To answer these questions and more, we need to turn to examples, see  Section~\ref{sec:gaussian_variance}.

\TODO{This bound might not hold in general, but it can along the trajectory. Can we prove it using an inductive method such as \citet[Appendix A]{lacoste2012simpler}?}

\section{Variance bound for centered gaussians}
\label{sec:gaussian_variance}
Sections on proximal point and mirror descent gave us the variance assumption 
\begin{align}
	\Cov(T(x_{n}) , \natp_{n+1}) = \Cov_X \bigg(T(X), \nabla \conj (\lr_n T(X) + (1-\lr_n)\mu_n)\bigg	) 
	\leq \lr_n M \; .
\end{align}
on the sufficient statistic and $\conj$. Unfortunately this bound is not trivially satisfied, even for simple exponential family members, such as centered gaussians $\cN(0,\sigma^2)$, when we are aiming to estimate the variance.
The density of a centered normal variable is
\begin{align}
	X \sim \cN(0,\sigma^2) \implies p(x) = \inv{\sqrt{2\pi \sigma^2}} e^{-\frac{x^2}{2 \sigma^2}} \; .
\end{align}
Using the $X^2$ as a sufficient statistic we get the variance as a mean parameter and minus half the precision as a natural parameter
\begin{align}
	\quad T(X)= X^2 \ ; \quad  \mu = \sigma^2 \ ; \quad   \natp = -\inv{2 \sigma^2} = -\inv{2 \mu}\; .
\end{align}
Now we can match the log-likelihood with the exponential family template to get the log-partition function.
\begin{align}
	-\log p(x) = \frac{x^2}{2 \sigma^2} - \half \log(2\inv{2\sigma^2}) + \half \log(2 \pi) 
	= -x^2 \natp + \logpart(\natp) \\
	\implies \logpart (\natp) = -\half \log(-\natp)  + \half \log(\pi)
\end{align}
We can then use the formula $\conj(\mu) = \mu \natp - \logpart(\natp)$ to get the entropy
\begin{align}
	\conj(\mu) = -\half \log(\mu) + \half \log\frac{\pi}{2} - \half \\
\end{align}
Remark that both the entropy and the log-partition have the same shape $z\mapsto -\inv{2z}$. Which yields the same shape of Bregman divergence, as visible below (all three lines are equal)
\begin{align}
	\KL( \sigma_n^2 ; \sigma_*^2) 
	&= \half \left ( \frac{\sigma_*^2}{ \sigma_n^2} - 1 - \log \frac{\sigma_*^2}{ \sigma_n^2} \right) \\
	\bregmanconj( \mu_*; \mu_n) 
	&= \half \left ( \frac{\mu_*}{ \mu_n} - 1 - \log  \frac{\mu_*}{ \mu_n} \right) \\
	\bregman( \natp_n; \natp_* ) 
	&=  \half \left ( \frac{ \natp_n}{\natp_*} - 1 - \log  \frac{ \natp_n}{\natp_*} \right) \; .
\end{align}
In other words, this is divergence measures the discrepancy between the ratio $\frac{ \natp_n}{\natp_*} =  \frac{\mu_*}{ \mu_n}  $ and $1$ via the function $\phi(z)= \half (z - 1 - \log(z))$.
\begin{align}
	\phi(z) := \half (z - 1 - \log(z)) 
	\quad \text{ and } \quad
	\frac{ \natp_n}{\natp_*} =  \frac{\mu_*}{ \mu_n}  \\
	\implies \bregman( \natp_n; \natp_* )   = \phi(\frac{ \natp_n}{\natp_*}) =  \phi(\frac{\mu_*}{ \mu_n})
\end{align}
The covariance  \eqref{eq:hypvariance} we want to bound is equal to
\begin{align}
	\half \expect{\frac{\sigma_*^2 - X^2 }{ X^2 \lr + (1-\lr) \sigma_0^2 }}
	\approx \lr (1-\lr)^{-2} \left (\frac{\sigma_*}{\sigma_0} \right)^4
\end{align}
using a first order approximation of the inverse, eg a first order approximation of $\nabla \conj$ that holds when $\lr X^2 \ll (1-\lr) \sigma_0^2$. This does not hold for $\sigma_0\ll 1$, which is precisely when the covariance  explodes, as illustrated in Figure \TODO{}. This is a problem, because we need this bound to hold  for all values of $\sigma_0$ in the support of $X^2$, meaning on the whole $[0,+\infty[$, which is impossible because $1/X^2$ diverges, so we need to process this integral differently.

@@

\paragraph{Big picture.}
Intuitively, we have summarize in the part in $\mathbb{P}(\norm{(\nabla A)^{-1}(\mu)} < t)$, the problem that arise with ${\cal B}_{A^*}(\mu^*; \mu)$ when $\mu$ is small. 
We could make this intuition more formal with
\begin{align}
   {\cal B}_{A^*}(\mu^*; \mu)
   &= 1_{\norm{\mu} \geq s} {\cal B}_{A^*}(\mu^*; \mu)
   + 1_{\norm{\mu} < s} {\cal B}_{A^*}(\mu^*; \mu)\\
   &\leq \sup_{\norm{\mu'} \geq s}(\norm{(\nabla A)^{-1}(\mu')})^2 \norm{\mu - \mu^*}^2
   + 1_{\norm{\mu} < s} {\cal B}_{A^*}(\mu^*; \mu).
\end{align}
Suppose the $\sup$ is finite, the first term is easy to control.
The second term is what we should take care of. We have
\begin{align}
    1_{\norm{\mu} < s} {\cal B}_{A^*}(\mu^*; \mu) 
    &=  1_{\norm{\mu} < s} (A^*(\mu^*) - A^*(\mu) - <\nabla A^*(\mu), \mu^* - \mu>)\\
    &\leq  1_{\norm{\mu} < s} (A^*(\mu^*) - \inf_{\norm{\mu'} < s} A^*(\mu') + \norm{\nabla A^*(\mu)} \sup_{\norm{\mu''} < s} \norm{\mu^* - \mu''}).
\end{align}
We know that, under some strict convexity condition, $\nabla A^* = (\nabla A)^{-1}$ (similar to Eq. (25) at time of writing). Therefore we really have to study, with harsh notation
\begin{align}
    \int_{0}^s \norm{(\nabla A)^{-1}(\mu)} d\mathbb{P}(\norm{\mu} < t)
    =\int_{0}^s \norm{(\nabla A)^{-1}(\mu)} \mathbb{P}(\norm{\mu} \in [t, t+dt]).
\end{align}
It seems to me that with the precedent development we were implicitly studying
\begin{align}
    \int_{0}^s \norm{(\nabla A)^{-1}(\mu)}^2 d\mathbb{P}(\norm{\mu} < t),
\end{align}
which of course is worse.
The study can be done precisely as we know the distribution of $\mu$.
In terms of rates, in my current view, this path is tight if and only if $<\nabla A^*(\mu), \mu^* - \mu>$ do behave like $\norm{(\nabla A)^{-1}(\mu)}$ towards $0$. In terms of constants, we have done pretty harsh majoration in order to get back to $c\norm{\mu-\mu^*}$ with constants $c$ that are clearly suboptimal.

@@

\section{Self-Concordance}

A big problem is that $\conj$ is seldom Lipschitz or smooth. For instance the log-partition function of a multivariate normal is 
\begin{align}
    \logpart(\eta, \Lambda) = \half \eta^\top \Lambda^{-1} \eta - \log\det(\Lambda)
\end{align}
which is defined on $\eta \in \real^d$ and $\Lambda\in \real^{d\times d}$ symmetric positive definite. 
It is not strongly convex, so $\conj$ is not smooth.

>>>>

\tableofcontents

@@

\subsection{Transferring SMD's covariance assumption to MAP estimates}
In our setting, 
\begin{align}
	\nabla f(\natp_t) - g_t =  T(X_{t+1})	- \E[T(X)]
\end{align}
so that $\sigma$ is really a bound on the covariance of the sufficient statistics with another variable
\begin{align}
	&\Cov( T(X_{t+1}), \natp_{t+1}| X_{1\dots t})
	\leq \lr_t\sigma^2 \\
	=& \Cov(T(X_{t+1}), \nabla\conj \left(\lr_t T(X_{t+1}) + (1-\lr_t ) \mu_t \right) | X_{1\dots t}) \;.
\end{align}
In other words, we need for all $\lr\in(0,\inv{n})$, and a certain subset of $\mu$,
\begin{align}
\label{eq:hypvariance}
	\boxed{
	\Cov_X \bigg(T(X), \nabla \conj (\lr T(X) + (1-\lr) \mu)\bigg	) 
	\leq \lr \sigma^2 \; .
	}
\end{align}

\paragraph{MLE Analysis.}
When we plug this bound into the expected suboptimality formula~\eqref{eq:bregmanMLE} with the maximum likelihood estimate $\hat \mu =  \inv{n}\sum_i T(X_i)$, we get
\begin{multline}
	\expect[X_{1\dots n}]{\bregmanconj ( \mu_* ;\hat \mu) }
	=\conj(\mu_*) \overbrace{- \E[\conj(\hat \mu)]}^{\leq - \conj(\E[\hat \mu])} \\
	+ \inv{n}\sum_i \E[\underbrace{\E[\langle  T(X_i) - \mu_*; \nabla\conj (\hat \mu) \rangle | X_j, j\neq i]}_{\leq \nicefrac{\sigma^2}{n} \eqref{eq:hypvariance}
	 	\; \text{with}\; \mu=\inv{n-1}\smallsum_{j\neq i} T(x_j)  } ]  \leq \frac{\sigma^2}{n}
\label{eq:MLE_bound}
\end{multline}
where we used the decomposition $\hat \mu = \inv{n}T(X_i) + (1-\inv{n}) \inv{n-1} \sum_{j\neq i}T(X_j)$ to apply~\eqref{eq:hypvariance}. 
In words, this variance assumption on $T(X)$ and $\conj$ immediately gives us a bound on the suboptimality.

This is too simple to be true.
Indeed we study assumption \eqref{eq:hypvariance} for the special case of the variance of a Gaussian in Section \ref{sec:gaussian_variance}.
In particular we show that it cannot hold for all possible values of $\mu$ within the double expectation in~\eqref{eq:MLE_bound}.
The mean parameter $\mu$ take values arbitrarily close from $0$, which makes the covariance explode.
\RLP{should we get a functional bound $\sigma^2(\mu)$ instead ? or split between $\mu$ smaller or larger than $\varepsilon$.}

\paragraph{MAP Analysis.}
We can also apply this result to the MAP estimate $\hat \mu = \frac{n_0 \mu_0 + \smallsum_i T(X_i)}{n_0+n}$ but I did not manage to reach a satisfying conclusion about the bias. 
Note $\tilde \mu = \E[\hat \mu] = \frac{n_0 \mu_0 + n \mu_*}{n_0+n}$ and $\gamma_0= \frac{n_0}{n_0+n}$ and the step-size $\gamma = \inv{n_0 + n}$ .
\begin{align}
	\expect[X_{1\dots n}]{\bregmanconj ( \mu_* ;\hat \mu) } 
	&\leq \conj(\mu_*) - \conj(\tilde \mu) + \gamma_0 \langle \mu_0 - \mu_* ; \E[\hat \natp] \rangle + (1 - \gamma_0) \gamma \sigma^2  \\
	& = \bregmanconj (\mu_* ; \tilde \mu) - \langle \tilde \mu - \mu_* ;    \tilde \natp - \E[\hat \natp] \rangle + \frac{n}{(n_0+n)^2}\sigma^2
\end{align}
so we recover the $O(n^{-1})$ for the variance and we are still looking for $O(n^{-2})$ rate for the bias. The Bregman term in the bias should be in this spirit, given a quadratic approximation
\begin{align}
	 \bregmanconj (\mu_* ; \tilde \mu)  
	 = \bregmanconj (\mu_* ; \mu_* + \gamma_0 (\mu_0 - \mu_*) ) \approx \half \gamma_0^2 \norm{\mu_0 - \mu_*}^2_{\mSigma_*^{-1}}
\end{align}
where $\mSigma_* = \Cov_{\mu_*}(T(X))$. 
The order of terms $\bregmanconj (\mu_* ; \tilde \mu)  \neq \bregmanconj (\tilde \mu ; \mu_* ) $ only affects third order terms.
That's for the bregman, but the scalar product is harder to bound. Hopefully it could be positive. If not we know that $\tilde \mu - \mu_* = \frac{n_0}{n_0+n} (\mu_0 - \mu_*) = O(n^{-1})$ , and we can hope that the same holds for $\tilde \natp - \E[\hat \natp] = \nabla \conj(\E[\hat \mu]) - \E[\nabla\conj (\hat \mu)]$, which kinda measures the non-linearity, or the curvature of $\conj$. Using a simple cauchy-schwartz, and the hessian of the conjugate, we might be able to get something.

@@

\paragraph{Bound on the trajectory.}
Bound \eqref{eq:hypvariance} might not hold in general, but it can along the trajectory. Can we prove it using an inductive method such as \citet[Appendix A]{lacoste2012simpler}?
No we cannot because $\mu$ takes every possible values as a random variable within the expectation.

\section{Expected Relative Smoothness}
To avoid gradient variance assumption, we can take inspiration from expected smoothness, which is a consequence of convexity and smoothness, and which yields a general analysis of SGD. Can we get something similar with relativity and SMD? 
The analog of gradient norm is the Bregman between iterates, similar to the Homeomorphic EM rate.

\section{Finding the variance of a Gaussian}
\label{sec:gaussian_variance}
Sections on proximal point and mirror descent gave us the variance assumption 
\begin{align}
	\Cov(T(x_{n}) , \natp_{n+1}) = \Cov_X \bigg(T(X), \nabla \conj (\lr_n T(X) + (1-\lr_n)\mu_n)\bigg	) 
	\leq \lr_n M \; .
\end{align}
on the sufficient statistic and $\conj$. Unfortunately this bound is not trivially satisfied, even for simple exponential family members, such as centered gaussians $\cN(0,\sigma^2)$, when we are aiming to estimate the variance.
The density of a centered normal variable is
\begin{align}
	X \sim \cN(0,\sigma^2) \implies p(x) = \inv{\sqrt{2\pi \sigma^2}} e^{-\frac{x^2}{2 \sigma^2}} \; .
\end{align}
Using the $X^2$ as a sufficient statistic we get the variance as a mean parameter and minus half the precision as a natural parameter
\begin{align}
	\quad T(X)= X^2 \ ; \quad  \mu = \sigma^2 \ ; \quad   \natp = -\inv{2 \sigma^2} = -\inv{2 \mu}\; .
\end{align}
Now we can match the log-likelihood with the exponential family template to get the log-partition function.
\begin{align}
	-\log p(x) = \frac{x^2}{2 \sigma^2} - \half \log(2\inv{2\sigma^2}) + \half \log(2 \pi) 
	= -x^2 \natp + \logpart(\natp) \\
	\implies \logpart (\natp) = -\half \log(-\natp)  + \half \log(\pi)
\end{align}
We can then use the formula $\conj(\mu) = \mu \natp - \logpart(\natp)$ to get the entropy
\begin{align}
	\conj(\mu) = -\half \log(\mu) + \half \log\frac{\pi}{2} - \half \\
\end{align}
Remark that both the entropy and the log-partition have the same shape $z\mapsto -\inv{2z}$. Which yields the same shape of Bregman divergence, as visible below (all three lines are equal)
\begin{align}
	\KL( \sigma_n^2 ; \sigma_*^2) 
	&= \half \left ( \frac{\sigma_*^2}{ \sigma_n^2} - 1 - \log \frac{\sigma_*^2}{ \sigma_n^2} \right) \\
	\bregmanconj( \mu_*; \mu_n) 
	&= \half \left ( \frac{\mu_*}{ \mu_n} - 1 - \log  \frac{\mu_*}{ \mu_n} \right) \\
	\bregman( \natp_n; \natp_* ) 
	&=  \half \left ( \frac{ \natp_n}{\natp_*} - 1 - \log  \frac{ \natp_n}{\natp_*} \right) \; .
\end{align}
In other words, this is divergence measures the discrepancy between the ratio $\frac{ \natp_n}{\natp_*} =  \frac{\mu_*}{ \mu_n}  $ and $1$ via the function $\phi(z)= \half (z - 1 - \log(z))$.
\begin{align}
	\phi(z) := \half (z - 1 - \log(z)) 
	\quad \text{ and } \quad
	\frac{ \natp_n}{\natp_*} =  \frac{\mu_*}{ \mu_n}  \\
	\implies \bregman( \natp_n; \natp_* )   = \phi(\frac{ \natp_n}{\natp_*}) =  \phi(\frac{\mu_*}{ \mu_n})
\end{align}
The covariance  \eqref{eq:hypvariance} we want to bound is equal to
\begin{align}
	\half \expect{\frac{\sigma_*^2 - X^2 }{ X^2 \lr + (1-\lr) \sigma_0^2 }}
	\approx \lr (1-\lr)^{-2} \left (\frac{\sigma_*}{\sigma_0} \right)^4
\end{align}
using a first order approximation of the inverse, eg a first order approximation of $\nabla \conj$ that holds when $\lr X^2 \ll (1-\lr) \sigma_0^2$. This does not hold for $\sigma_0\ll 1$, which is precisely when the covariance  explodes, as illustrated in Figure \TODO{}. This is a problem, because we need this bound to hold  for all values of $\sigma_0$ in the support of $X^2$, meaning on the whole $[0,+\infty[$, which is impossible because $1/X^2$ diverges, so we need to process this integral differently.

@@

\subsection{Big picture.}
Intuitively, we have summarize in the part in $\mathbb{P}(\norm{(\nabla A)^{-1}(\mu)} < t)$, the problem that arise with ${\cal B}_{A^*}(\mu^*; \mu)$ when $\mu$ is small. 
We could make this intuition more formal with
\begin{align}
   {\cal B}_{A^*}(\mu^*; \mu)
   &= 1_{\norm{\mu} \geq s} {\cal B}_{A^*}(\mu^*; \mu)
   + 1_{\norm{\mu} < s} {\cal B}_{A^*}(\mu^*; \mu)\\
   &\leq \sup_{\norm{\mu'} \geq s}(\norm{(\nabla A)^{-1}(\mu')})^2 \norm{\mu - \mu^*}^2
   + 1_{\norm{\mu} < s} {\cal B}_{A^*}(\mu^*; \mu).
\end{align}
Suppose the $\sup$ is finite, the first term is easy to control.
The second term is what we should take care of. We have
\begin{align}
    1_{\norm{\mu} < s} {\cal B}_{A^*}(\mu^*; \mu) 
    &=  1_{\norm{\mu} < s} (A^*(\mu^*) - A^*(\mu) - <\nabla A^*(\mu), \mu^* - \mu>)\\
    &\leq  1_{\norm{\mu} < s} (A^*(\mu^*) - \inf_{\norm{\mu'} < s} A^*(\mu') + \norm{\nabla A^*(\mu)} \sup_{\norm{\mu''} < s} \norm{\mu^* - \mu''}).
\end{align}
We know that, under some strict convexity condition, $\nabla A^* = (\nabla A)^{-1}$ (similar to Eq. (25) at time of writing). Therefore we really have to study, with harsh notation
\begin{align}
    \int_{0}^s \norm{(\nabla A)^{-1}(\mu)} d\mathbb{P}(\norm{\mu} < t)
    =\int_{0}^s \norm{(\nabla A)^{-1}(\mu)} \mathbb{P}(\norm{\mu} \in [t, t+dt]).
\end{align}
It seems to me that with the precedent development we were implicitly studying
\begin{align}
    \int_{0}^s \norm{(\nabla A)^{-1}(\mu)}^2 d\mathbb{P}(\norm{\mu} < t),
\end{align}
which of course is worse.
The study can be done precisely as we know the distribution of $\mu$.
In terms of rates, in my current view, this path is tight if and only if $<\nabla A^*(\mu), \mu^* - \mu>$ do behave like $\norm{(\nabla A)^{-1}(\mu)}$ towards $0$. In terms of constants, we have done pretty harsh majoration in order to get back to $c\norm{\mu-\mu^*}$ with constants $c$ that are clearly suboptimal.

@@

\section{Self-Concordance}
\subsection{Motivation}
A big problem is that $\conj$ is seldom Lipschitz or smooth. For instance the log-partition function of a multivariate normal is 
\begin{align}
    \logpart(\eta, \Lambda) = \half \eta^\top \Lambda^{-1} \eta - \log\det(\Lambda)
\end{align}
which is defined on $\eta \in \real^d$ and $\Lambda\in \real^{d\times d}$ symmetric positive definite. 
It is not strongly convex, so $\conj$ is not smooth.
